---
title: "STAA566_Schaefer_hw4"
author: "Paul Schaefer"
format: html
editor: visual
---

```{r}
# Inport libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(reactable)
```


## Data import and cleaning
```{r message = FALSE}
gas <- read_csv("C:\\Users\\pscha\\Documents\\_School\\2022_Fall_Classes\\STAA_566_DataViz\\natgas.csv")
gas <- gas[complete.cases(gas[ , 8]),] # get rid on NAs in Capacity
cap <- sum(gas$Capacity)
rank <- c(1:nrow(gas))
gas <- gas %>%
  mutate(percent=Capacity/cap) #calculate percentage of total
gas <- gas[order(-gas$percent),]
gas <- gas %>%
  mutate(cum_per=cumsum(percent),rank = rank)%>%# calculate cumulative percentage
  mutate(percentile = case_when(cum_per<=.05 ~ '5th',
                                cum_per<=.1 & cum_per > .05 ~ '10th',
                                cum_per<=.2 & cum_per > .1 ~ '20th',
                                cum_per > .2 ~ 'higher'))#%>%
  #mutate(display = case_when(cum_per<=.05 ~ cell_spec,
  #                              cum_per<=.1 & cum_per > .05 ~ '10',
  #                              cum_per<=.2 & cum_per > .1 ~ '20',
  #                              cum_per > .2 ~ 'higher')))

head(gas)
```
## Make the table
```{r}
gas_table2 <- gas %>%
  select(rank,Name, Owner, State, Capacity,percentile) %>%
  reactable(
    defaultColDef = colDef(align = 'center'),
    columns = list(
      rank = colDef(name = "Rank", align = 'right'),
      Name = colDef(name = "Name", align = 'left'),
      State = colDef(name = "State", align = 'left'),
      Owner = colDef(name = "Owner", align = 'left'),
      Capacity = colDef(name = "Capacity (M cu. ft. per day)", align = 'right'),
      percentile = colDef(name = 'Percentile'), align = 'right'),
    style = list(fontSize = "1rem"))
gas_table2
```

This table is based on a work project a few years back when we attempted to determine whether there were parts of the US energy infrastructure that contributed a 'dispoportionate' amount to the overall capacity. We wanted to know if there were a 'relatively small' number of production facilities that produced a 'relatively large' percentage of different outputs. That project involved Excel, which meant the tables were dreadful, so we ended up putting the results on a map instead. This is an attempt to produce a table that gives the same information.

I limited the coulmns to the ones that we would have used for the project: the name, owner, capacity, and the percentile that the plant fell in. Due to the size of the data, I wanted tabs, so I used the reactable() package for the table.  I had some code problems with the percentiles - there are some plants with the same capacity, so they assigned percentiles by the order they were originally in, but sorting on 'Capacity' will slightly mis-order them.

The data source is a .csv downloaded from the US Energy Information Agency website: https://www.eia.gov/naturalgas/ngqs/#?report=RP9&year1=2017&year2=2017&company=Name 
The data is based on 2017 reporting and was released in 2019.
