---
title: "STAA566_Schaefer_hw4"
author: "Paul Schaefer"
format: html
editor: visual
---

```{r}
# Inport libraries
library(tidyverse)
```


## Data import and cleaning
```{r message = FALSE}
gas <- read_csv("C:\\Users\\pscha\\Documents\\_School\\2022_Fall_Classes\\STAA_566_DataViz\\natgas.csv")
gas <- gas[complete.cases(gas[ , 8]),]
cap <- sum(gas$Capacity)
typeof(gas$Capacity)
typeof(cap)
gas <- gas %>%
  mutate(percent=Capacity/cap)
gas <- gas[order(-gas$percent),]
gas <- gas %>%
  mutate(cum_per=cumsum(percent))
head(gas)
```
This table is based on a work project a few years back when we attempted to determine whether there were parts of the US energy infrastructure that contributed a 'dispoportionate' amount to the overall capacity. We wanted to know if there were a 'relatively small' number of production facilities that produced a 'relatively large' percentage of different outputs. That project involved Excel, which meant the tables were dreadful, so we ended up putting the results on a map instead. This is an attempt to produce a table that gives the same information.

The data source is a .csv downloaded from the US Energy Information Agency website: https://www.eia.gov/naturalgas/ngqs/#?report=RP9&year1=2017&year2=2017&company=Name 
The data is based on 2017 reporting and was released in 2019.
