---
title: "STAA566_Schaefer_hw4"
author: "Paul Schaefer"
format: html
editor: visual
---

```{r}
# Inport libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(reactable)
```


## Data import and cleaning
```{r message = FALSE}
gas <- read_csv("C:\\Users\\pscha\\Documents\\_School\\2022_Fall_Classes\\STAA_566_DataViz\\natgas.csv")
gas <- gas[complete.cases(gas[ , 8]),] # get rid on NAs in Capacity
cap <- sum(gas$Capacity)
rank <- c(1:nrow(gas))
gas <- gas %>%
  mutate(percent=Capacity/cap) #calculate percentage of total
gas <- gas[order(-gas$percent),]
gas <- gas %>%
  mutate(cum_per=cumsum(percent),rank = rank)%>%# calculate cumulative percentage
  mutate(percentile = case_when(cum_per<=.05 ~ '5th',
                                cum_per<=.1 & cum_per > .05 ~ '10th',
                                cum_per<=.2 & cum_per > .1 ~ '20th',
                                cum_per > .2 ~ 'higher'))#%>%
  #mutate(display = case_when(cum_per<=.05 ~ cell_spec,
  #                              cum_per<=.1 & cum_per > .05 ~ '10',
  #                              cum_per<=.2 & cum_per > .1 ~ '20',
  #                              cum_per > .2 ~ 'higher')))

head(gas)
```
## Make the table
```{r}
# make table
gas_table <- gas %>%
  select(Name, Owner,Capacity,percent,cum_per) %>%
  kbl(escape = FALSE,
      col.names = c("Name",
                    "Owner",
                    "Capacity",
                    "Percentage",
                    "Cumulative Percentage"
                    ),
      align = c("l","l","r","r"),
      digits = 4) %>%
  #add_header_above(c(" ", "14-day average" = 4, " ", "14-day average per 100k" = 2)) %>%
  kable_styling(font_size=18, fixed_thead = TRUE)%>%
  scroll_box(width = "100%", height = "700px")
gas_table
```
## Make another table
```{r}
gas_table2 <- gas %>%
  select(rank,Name, Owner, Capacity, percent, cum_per,percentile) %>%
  reactable(
    columns = list(
      rank = colDef(name = "Rank"),
      Name = colDef(name = "Name"),
      Owner = colDef(name = "Owner"),
      Capacity = colDef(name = "Capacity (M cu. ft. per day)"),
      percent= colDef(name = "Percent", format=colFormat(digits=3)),
      cum_per= colDef(name = "Cumulative Percent", format=colFormat(digits=3)),
      percentile = colDef(name = 'Percentile')),
    style = list(fontSize = "1rem"))
gas_table2
```

This table is based on a work project a few years back when we attempted to determine whether there were parts of the US energy infrastructure that contributed a 'dispoportionate' amount to the overall capacity. We wanted to know if there were a 'relatively small' number of production facilities that produced a 'relatively large' percentage of different outputs. That project involved Excel, which meant the tables were dreadful, so we ended up putting the results on a map instead. This is an attempt to produce a table that gives the same information.

The data source is a .csv downloaded from the US Energy Information Agency website: https://www.eia.gov/naturalgas/ngqs/#?report=RP9&year1=2017&year2=2017&company=Name 
The data is based on 2017 reporting and was released in 2019.
